
# Bivariate Analysis

## Quantitative VS Outcome

Outcome = `DxCancer`

### Vizualisation

#### Boxplots by diagnosis

```{r Quantitative VS Outcome plots ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

quantitative <- colnames(data[colnames(data) %in% quantitative_variable_names])

# independant scatterplots for each pairs of variables
plot_list <- list()  # list to stock graph

for (i in 1:(length(quantitative)-1)) {
  
  var_x <- Outcome  # Outcome
  var_y <- quantitative[i]  # quantitative var

  if (var_x != var_y) {
    # boxplot creation
    p <- ggplot(data, aes(x = .data[[var_x]], y = .data[[var_y]], fill = .data[[var_x]])) +
      geom_boxplot() +
      scale_fill_manual(values = c("TRUE" = "#83b9cb", "FALSE" = "#f49691")) +  # Colorer TRUE en bleu, FALSE en rouge
      labs(title = paste("Boxplot of", var_y, "by", var_x)) +
      theme_minimal() +
      theme(plot.title = element_text(size = 10),
            plot.margin = unit(c(1, 1, 3, 1), "lines"),
            legend.position = "none")  

    # add plot 
    plot_list[[paste(var_y, var_x, sep = "_by_")]] <- p
  }
}

# display graph 
aes_columns_display(plot_list, 2)
```



#### Density plots by diagnosis

```{r Quantitative VS Outcome density ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

density_plot_list <- list()    # list to stock graph

for (i in 1:length(quantitative)) {
  
    var_x <- quantitative[[i]] 
    var_y <- Outcome  

    if (!identical(var_x, var_y)) {  # var != outcome
      
        # graph
        p <- ggplot(data, aes_string(x = var_x, fill = var_y, colour = var_y)) +
                  geom_density(alpha = 0.25)+
          geom_rug() +
          labs(title = paste("Density plot of", var_x, "by Cancer diag"),
               x = names(quantitative)[i],
               fill = "Cancer diag",
               colour = "Cancer diag")+
      theme_minimal() + 
      theme(plot.title = element_text(size = 10))
        
   density_plot_list[[paste(var_y, var_x, sep = "_by_")]] <- p
    }
    else {
       
    }
  }


# display graph 
aes_columns_display(density_plot_list, 2)
```

### Statistics

This statistical section will allow us to observe whether there is a 
significant difference between patients diagnosed with cervical cancer and 
undiagnosed patients according to the different quantitative variables in the study. 
We begin by testing the normality of the variables to find out the appropriate statistical test to use. 

#### hypothesis verification
##### Normality test for continuous variables
###### Shapiro
```{r Shapiro test ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}


for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]
res <- shapiro.test(data[[var]])

  cat("\nShapiro Test Result for", var, "distribution :\n")
  print(res)
  cat("\n")

}

```

###### QQplot
```{r qq visualisation ,echo = FALSE, message=FALSE, warning=FALSE, results='hide', fig.height = 4, fig.width=5}


for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]

p<-qqnorm(data[[var]], main = NULL)
  qqline(data[[var]])
  title(main = paste(var))
}

```


##### Variances test for continuous variables
```{r variance levene Test,,echo = FALSE, message=FALSE, warning=FALSE, results='hide', fig.height = 4, fig.width=5}


for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]
res <- leveneTest(data[[var]] ~ DxCancer, data = data)

  cat("\nLevene Test Result for", var, "variance :\n")
  print(res)
  cat("\n")

}
```

```{r variance fligner Test,,echo = FALSE, message=FALSE, warning=FALSE, results='hide', fig.height = 4, fig.width=5}


for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]
res <- fligner.test(data[[var]] ~ DxCancer, data = data)

  cat("\nFligner Test Result for", var, "variance :\n")
  print(res)
  cat("\n")

}
```

None of the distributions is normal. The Wilcoxon test will be used.


#### test choice

The wilcoxon test will allow us to establish whether there is a significant difference between 
patients diagnosed with cancer and those not diagnosed for the different quantitative variables in the study.

```{r wilcoxon tests ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 10}

wilcox_list <- list()

for (i in 1:length(continuous_variable_names)) {
var <- continuous_variable_names[[i]]

res <- wilcox.test(data[[var]]~data[[Outcome]],conf.int = TRUE)
wilcox_list[[paste(var)]] <- res
}

wilcox_list
```

The statistical test assumes that there is a statistically significant difference between the groups for age, IUDYears (Years of IUD use) (p-value < 0.05)  

#### Size effect :

```{r cohen d , echo = FALSE, message=FALSE, warning=FALSE, fig.height = 10}

for (i in 1:length(continuous_variable_names)) {
  var <- continuous_variable_names[i]
  
  group_true <- data[[var]][data$DxCancer == "TRUE"]
  group_false <- data[[var]][data$DxCancer == "FALSE"]
  
  # Cohen's d
  res <- cohen.d(group_true, group_false)
  
  # display results
  cat("\nCohen's d Result for", var, "by Cancer diagnosis is:\n")
  print(res)
  cat("\n")
}

```


## Qualitative VS Outcome

### Vizualisation and statistics

The Pearson's Chi-squared test will allow us to establish whether there is a significant difference between 
patients diagnosed with cancer and those not diagnosed for the different qualitative variables in the study.

```{r Qualitative VS Outcome plots,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

qualitative <- colnames(data[colnames(data) %in% qualitative_variable_names])


plot_list <- list()


for (i in 1:length(qualitative)) {
  
    var_x <- qualitative[i]  
    var_y <- Outcome        

    if (var_x != var_y){

      p <- ggplot(data, aes(x = .data[[var_x]], fill = factor(.data[[var_y]]))) +
        geom_bar(position = "fill") +
        scale_y_continuous(labels = scales::percent_format()) +  
        labs(title = paste("Proportion of", var_y, "by", var_x), 
             x = var_x, y = "Proportion", fill = var_y) +
        theme_minimal() +
        theme(plot.title = element_text(size = 10),
            plot.margin = unit(c(1, 1, 3, 1), "lines"))
      

      plot_list[[paste(var_x, var_y, sep = "_by_")]] <- p
    }
}

```


```{r Chi2 tests ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4}

chi2_list <- list()

for (i in 1:length(qualitative)) {
  
    var_x <- qualitative[i] 
    var_y <- Outcome

table_res <- table(data[[var_x]], data[[var_y]])
res <- chisq.test(table_res)

 chi2_list[[paste(var_x, var_y, sep = "_by_")]] <- res
}

```


```{r Qualitative display ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4, fig.width=5}

for (i in 1:length(qualitative)) {

  var_x <- qualitative[i]  
  var_y <- Outcome

  if (var_x != var_y){

  p <- plot_list[[paste(var_x, var_y, sep = "_by_")]]
  print(p)
  

  res <- chi2_list[[paste(var_x, var_y, sep = "_by_")]]
  cat("\nChi-Square Test Result for", var_x, "and", var_y, "association :\n")
  print(res)
  cat("\n")
  cat("\n------------- Effect size : ------------\n")
  cat("\nCramer V Test Result for", var_x, "and", var_y, "association :\n")
  cramerV(res$observed,ci = TRUE)
  print(cramerV(res$observed,ci = TRUE))


}
}
```




The statistical test assumes that there is a statistically significant difference between the groups for IUD (IUD use), DxHPV, Dx, Hinselmann, Schiller, Cytology, Biopsy (p-value < 0.05)  

#### Odd Ratios :

**Table 7 : Odds Ratios**

```{r Odds Ratios univariate ,echo = FALSE, message=FALSE, warning=FALSE, fig.height = 10}

library(finalfit)
res_glm_uni <- data%>%
    glmuni("DxCancer", qualitative[-15]) %>% 
    fit2df(estimate_suffix=" (univarié)")

kable(res_glm_uni,row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))

```