

```{r functions, echo = FALSE, message=FALSE, warning=FALSE}

#'check_event_count(), 
#'check the event number in the targeted variable (DxCancer) for each variables.
#' 
#' data : dataframe of all tested variables
#' target_var : targeted variable with events (DxCancer)
#' quantitative_variable_list : list
#' qualitative_variable_list : list
#' threshold : numeric

check_event_count <- function(data, target_var, quantitative_variable_list, qualitative_variable_list, threshold = 10) {

  result <- list()

#for all variables
  for (var in names(data)) {
  # if var is a quantitative variable 
    if (var %in% quantitative_variable_list) {
      #check nb of events in target variable
      count <- sum(data[[target_var]] == 1 & !is.na(data[[var]]), na.rm = TRUE)
      #if nb events is inf to 10 capture the name of the variable that deosn't get conditions
      if(count<10){
        result[[var]] <- var
        result[[var]][2] <- count
        }
    } 
  # if var is a qualitative variable
    if (var %in% qualitative_variable_list) {
      for (i in 1:length(unique(data[[var]]))){
      #check nb of events in target variable for all levels of the variable var
      factor_counts <- table(data[[var]], data[[target_var]])[ i, "TRUE"]
      #capture the name of the variable that deosn't get conditions
      if(factor_counts<10){
        result[[var]] <- var
        result[[var]][2] <- paste("level :",unique(data[[var]])[i])
        result[[var]][3] <-  paste("event(s) :",factor_counts)
        }
    } 

    } else { }
  }
  return(result)
  
} 


#' error_rate(), 
#' calculate the error rate
#' 
#' model : object from glm()
#' data : dataframe 
#' data_outcome : data$outcome

error_rate <- function(model,data,data_outcome){
  #proba beeing positive
  pplus <- predict(model,newdata=data,type="response")
  #get prediction
  prediction <- as.factor(ifelse(pplus > 0.5,"positive","negative"))
  #confusion matrix
  mc <- table(data_outcome,prediction)
  cat("Confusion matrix :\n")
  print(mc)
  #error rate
  err.rate <- (mc[2,1]+mc[1,2])/sum(mc)
  cat("\nError rate : ")
  cat(err.rate)
}

```

## Explicative

### Logistic Regression model

```{r set train/test list, echo = FALSE, message=FALSE, warning=FALSE,fig.height=7}

#reproducibility
set.seed(123)

#'_____________________________
#' 
#' SET TRAIN/TEST LIST   ----
#'_____________________________
#'

train_indices <- sample(1:nrow(data_dummy_post_colinear_analysis), size = 0.8 * nrow(data_dummy_post_colinear_analysis))

# train/test creation (comment if list already exist)
train_data <- data_dummy_post_colinear_analysis[train_indices, ]  # 80% training set
test_data <- data_dummy_post_colinear_analysis[-train_indices, ]  # 20% testing set

#creat csv for reproducibility (comment if list already exist)
write.csv(train_data, "train_data.csv", row.names = FALSE)
write.csv(test_data, "test_data.csv", row.names = FALSE)

train_data = read.csv("train_data.csv")
test_data = read.csv("test_data.csv")

#check point distribution proportion
p1 <- prop.table(table(data_dummy_post_colinear_analysis[[Outcome]]))
p2 <- prop.table(table(train_data[[Outcome]]))
p3 <- prop.table(table(test_data[[Outcome]]))

```



#### Modelisation : Model1

```{r full logistic model modelisation, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#'____________________________________
#' 
#' MODELISATION Model1
#'____________________________________

# Parameters estimation :
logistic_model_all <- glm(DxCancer ~ ., family = binomial(link = 'logit'), data = train_data)

# Description :
summary(logistic_model_all)
```

______________________________________________________________________________

**Hypothesis**

- Independence
- No colinearity 
- Sufficient sample size
- No extrem values for residue   

>https://pmc.ncbi.nlm.nih.gov/articles/PMC8710907/
______________________________________________________________________________

***Independence :***  

All the observations are independent. No time series or repeated measurements

***Colinearity check :***  
articles : colinearité

```{r check_collinearity, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
check_collinearity(logistic_model_all)
```

***Size sample check :***  
articles : 10 evenement pas feature
```{r check_event_count, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
event_table <- as.data.frame(check_event_count(data_dummy_post_colinear_analysis,"DxCancer",quantitative_variable_names,qualitative_variable_names))
event_table
```

***Extrem values for residue check :***  

```{r check_residue, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
par(mfrow = c(2, 2))
plot(logistic_model_all)
check_model(logistic_model_all)
```

Extremely large logits (such as −150) can occur for several reasons:

- Perfectly separable data:  
The predictors allow for almost perfect separation, leading the model to produce probabilities 
𝑝 that are extremely close to 0 or 1.

- Outliers or high leverage:  
One or more atypical observations exert a strong influence on the model.

- Multicollinearity:  
If the predictors are highly correlated, it can destabilize the estimation of the coefficients.

#### Statistics evaluation : Model1


**Likelihood ratio test :** 

*H0* : no significant difference between model with predictors and without predictors (B0 only)  
*H1* : significant difference between model with predictors and without predictors
pvalue < 0.05 H0 rejected : at least one predictor is significantly related to the dependent variable

```{r full logistic model Likelihood ratio test, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#'________________________
#' 
#'STATISTIC EVALUATION 
#'________________________

#chi-2 global significativity test
chi2 <- logistic_model_all$null.deviance - logistic_model_all$deviance
#ddl 
ddl <- logistic_model_all$df.null - logistic_model_all$df.residual
#p-value
pvalue <- pchisq(chi2,ddl,lower.tail = FALSE)

```

chi-2 global significativity test : `r chi2`  
ddl : `r ddl`  
pvalue : `r pvalue`  

```{r full logistic model R2, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}

#-- R2 McFadden
LLa <- logistic_model_all$deviance/(-2)
LL0 <- logistic_model_all$null.deviance/(-2)
R2MF <- 1.0 - LLa/LL0

#-- Cox-Snell
La <- exp(LLa)
L0 <- exp(LL0)
R2CS <- 1.0 - (L0/La)^(2/nrow(train_data))

#-- Nagelkerke
R2N <- R2CS/(1.0 - L0^(2/nrow(train_data)))

dfR2 <- data.frame("R2 McFadden" = R2MF, "R2 Cox-Snell" = R2CS, "R2 Nagelkerke" = R2N )
dfR2
```



**Odds Ratios :** 

```{r full logistic model OR, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
tbl_or <- tbl_regression(
  logistic_model_all,
  exponentiate = TRUE
) |> 
as_gt() |> 
  gt::tab_style(
    style = list(
      gt::cell_text(size = px(10))  # Réduction de la taille de la police
    ),
    locations = list(
      gt::cells_body(),              # Corps du tableau
      gt::cells_column_labels(),  # En-têtes de colonnes
      gt::cells_title()        
    )
  )
tbl_or
```

We can note extreme values and unsuable confidence interval. (pb : colinearity, overfitting, imbalance)  
(pas de surdispersion donc pas besoin de passer en quasi binomial)  
  
Remarques : le test de Wald sert à évaluer la significativité d’un prédicteur, tandis que le rapport de cotes mesure la force de l’effet.

**Variable importance :** 

```{r full logistic model VarImp, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#' Variable importance :
varImp(logistic_model_all)
```

proba very low and similar to each other the model possibly overlearns  


#### Predictive evaluation : Model1

```{r full logistic model prediction mod1, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#'________________________
#' 
#'PREDICTIVE EVALUATION 
#'________________________

pred1 <- predict(logistic_model_all,newdata=test_data,type="response")
pred1_class <- as.factor(ifelse(pred1 > 0.5,TRUE,FALSE))

error_rate(logistic_model_all,test_data,test_data$DxCancer)
```
Probability predicts too well (little diversity in the data, very few cases and/or collinearity +++)  

**Confusion matrix :** 

```{r full logistic model conf matrix mod1, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}

caret::confusionMatrix(data=pred1_class,reference=(as.factor(test_data$DxCancer)),positive="TRUE")
# /!\ ne prend que factor pas de logical !
```

**Hosmer et Lemeshow test :** 

*H0* : no significant difference between predicted data and observed data   
*H1* : significant difference between predicted data and observed data  
pvalue < 0.05 H0 rejected  

```{r full logistic model HosmerLemeshow test, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#Le test de Hosmer et Lemeshow (1989) 
print(generalhoslem::logitgof(test_data$DxCancer,pred1))
#print(hoslem.test(test_data$DxCancer, pred1, g = 3))
```

Remarque : negative ddl so, not enough data to properly divide your sample into three groups.

**ROC curve model1 prediction:** 

```{r full logistic model ROC, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
# ROC curve
pred <- ROCR::prediction(pred1,test_data$DxCancer)
graphs_roc <- ROCR::performance(pred,measure="tpr",x.measure="fpr")
ROCR::plot(graphs_roc,xlab="TFP",ylab="TFP",col="darkblue")
abline(a=0,b=1)
```

```{r}
#test 1 

#creation de meta-niveau 

#codage imbriqué pour indiquer une gradation

# attention, codage 1, 2, 3 
#niv1 <- ifelse(donnees$ALCOOL >= 1, 1, 0)
#niv2 <- ifelse(donnees$ALCOOL >= 6, 1, 0)
#niv3 <- ifelse(donnees$ALCOOL >= 11, 1, 0)

# régression avec codage imbriqué
#print(summary(glm(donnees$RONFLE ~ niv1+niv2+niv3,family=binomial)))

```

```{r}
#test 1 

#creation de meta-niveau 

#codage imbriqué pour indiquer une gradation

# attention, codage 1, 2, 3 
#niv1 <- ifelse(donnees$ALCOOL >= 1, 1, 0)
#niv2 <- ifelse(donnees$ALCOOL >= 6, 1, 0)
#niv3 <- ifelse(donnees$ALCOOL >= 11, 1, 0)

# régression avec codage imbriqué
#print(summary(glm(donnees$RONFLE ~ niv1+niv2+niv3,family=binomial)))

```

#### Evaluation des variables

```{r logistic model adjustment, echo = FALSE, message=FALSE, warning=FALSE,fig.height=7}
#'________________________
#' 
#' SELECTION DES VARIABLES
#'________________________

#selection des variables par p-value
#selection des variables apres anova ?
#selection des variables apres stepAIC() (MASS library) both forward backward
```

##### backward_model_AIC :

```{r logistic model AIC modelisation, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#'____________________
#' 
#' Variable selection “backward”  - AIC criterion ----
#'____________________

backward_model_AIC <- logistic_model_all %>%
                  stepAIC(direction = "backward")
#summary(backward_model_AIC)
```

```{r logistic model AIC prediction, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
pred2 <- predict(backward_model_AIC,newdata=test_data,type="response")
```

```{r logistic model AIC evaluation, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
error_rate(backward_model_AIC,test_data,test_data$DxCancer)
backward_model_AIC$anova
plot(0:(nrow(backward_model_AIC$anova)-1),backward_model_AIC$anova[,"AIC"],type="b",xlab="# de var. retirées",ylab="AIC",main="Sélection backward (AIC)")
```

**ROC curve for backward_model_AIC prediction:** 

```{r logistic model AIC ROC , echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
# ROC curve
pred <- ROCR::prediction(pred2,test_data$DxCancer)
graphs_roc <- ROCR::performance(pred,measure="tpr",x.measure="fpr")
ROCR::plot(graphs_roc,xlab="TFP",ylab="TFP",col="darkblue")
abline(a=0,b=1)
```

##### backward_model_BIC : 

```{r logistic model BIC modelisation, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
#'____________________
#' 
#' Variable selection “backward”  - BIC criterion ----
#'____________________

backward_model_BIC<-logistic_model_all %>%
                  stepAIC(direction = "backward",k = log(nrow(train_data)))
#summary(backward_model_BIC)
```

```{r logistic model BIC prediction, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
pred3 <- predict(backward_model_BIC,newdata=test_data,type="response")
```

```{r logistic model BIC evaluation, echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
error_rate(backward_model_BIC,test_data, test_data$DxCancer)
backward_model_BIC$anova
plot(0:(nrow(backward_model_BIC$anova)-1),backward_model_BIC$anova[,"AIC"],type="b",xlab="# de var. retirées",ylab="BIC",main="Sélection backward (BIC)")
#selection par lasso
```

**ROC curve for backward_model_BIC prediction:** 

```{r logistic model BIC ROC , echo = TRUE, message=FALSE, warning=FALSE,fig.height=7}
# ROC curve
pred <- ROCR::prediction(pred3,test_data$DxCancer)
graphs_roc <- ROCR::performance(pred,measure="tpr",x.measure="fpr")
ROCR::plot(graphs_roc,xlab="TFP",ylab="TFP",col="darkblue")
abline(a=0,b=1)
```

 cet article : https://www.researchgate.net/publication/325129765_Supervised_deep_learning_embeddings_for_the_prediction_of_cervical_cancer_diagnosis






