# Multivariate Analysis

To improve the results of the multivariate analysis, data has been normalised.  

>`scale()` is used from the heatmaply library. description: `scale()` use Z-score.  
This method is robust to extreme values and retains information on the variance. 

## Descriptive 

### Standardisation & Dummification

```{r correlation matrix function, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}

#'_________________________________________
#' 
#' Normalised (quanti) data correlation matrix   ----
#'_________________________________________


#cor() This function filters out any non-numerical variables and provides correlations only between quantitative variables. 
#Best for datasets with only a few variables. The correlation matrix is returned (with class matrix).

get_correlation_matrix <- function(data, Outcome = NULL, use = "pairwise.complete.obs") {

  # Select columns based on the presence of ID and Outcome
  if ("ID" %in% colnames(data)) {
    if (!is.null(Outcome)) {
      cor_matrix <- cor(data %>% dplyr::select(-ID, -Outcome), use = use)
    } else {
      cor_matrix <- cor(data %>% dplyr::select(-ID), use = use)
    }
  } else {
    if (!is.null(Outcome)) {
      cor_matrix <- cor(data %>% dplyr::select(-Outcome), use = use)
    } else {
      cor_matrix <- cor(data, use = use)
    }
  }

  # Create the heatmap to visualize the correlation matrix
  ggplot(data = as.data.frame(as.table(cor_matrix)), aes(Var1, Var2, fill = Freq)) +
    geom_tile() +
    geom_text(aes(label = round(Freq, 2)), color = "black", size = 3) +
    scale_fill_gradient2(low = "#3838ff", high = "#ff4444", mid = "#ffffff", midpoint = 0, limit = c(-1, 1), space = "Lab", name = "Correlation") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    coord_fixed() +
    labs(title = "", x = "", y = "")
}

```

**Figure 5 : Heatmap on initial data**

```{r headmap on data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}

#'__________________________________
#' 
#' Initial data HEADMAP   ----
#'__________________________________

headmap<- heatmaply(
  data, 
  xlab = "Features",
  ylab = "Patients", 
  main = "data headmap"
)
```


```{r normalisation, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 4} 

#'______________________________________________
#' 
#' STANDARDISATION (Z-score using scale())  ----
#'______________________________________________

#remove ID from the selection 
normalised_data <- data%>% dplyr::select(-ID)
#only quantitative selection 
quantitative_names<-colnames(normalised_data) %in% quantitative_variable_names
#standardisation with scale() fonction
normalised_data[quantitative_names] <- scale(normalised_data[quantitative_names])

```


```{r create dummy variables, echo = FALSE, message=FALSE, warning=FALSE,fig.height=7}

#'__________________________________________________
#' 
#' SET DUMMY variables on normalised dataset    ----
#'__________________________________________________
#'

data_dummy <- normalised_data
#excpet DxCancer to compare on patient with/without cancer diag
qualitative_without_Dxcancer <- setdiff(qualitative, "DxCancer")
#Dummification
data_dummy[,qualitative_without_Dxcancer] <- lapply(normalised_data[, qualitative_without_Dxcancer], as.numeric)

```

**Figure 6 : Correlation matrix on standardised and dummy data**

```{r correlation matrix on dummy data, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}

#'_________________________________________
#' 
#' dummy data correlation matrix   ----
#'_________________________________________

#outcome = "DxCancer"
correlation_matrix<-get_correlation_matrix(data_dummy,"DxCancer")
#correlation_matrix$data est de type facteur, alors lorsqu’elle est manipulée, les valeurs renvoyées sont les indices des niveaux du facteur, pas les noms eux-mêmes.
correlation_matrix$data$Var1 <- as.character(correlation_matrix$data$Var1)
colinear_variables_list<-unique(correlation_matrix$data$Var1[correlation_matrix$data$Freq != 1 & correlation_matrix$data$Freq >= 0.8])
correlation_matrix
```

On this figure, the values of the Pearson 
correlation coefficients (noted α) are indicated for each pair of variables.  

There are many variables that seem to have a linear relationship in the data frame 

**Figure 6a : High correlation matrix (<0.8)**
```{r colinearity treatment, echo = FALSE, message=FALSE, warning=FALSE,fig.height=7}

#'____________________
#' 
#' COLINEARITY  ----
#'____________________

#etude de la colinearité

p<-get_correlation_matrix(data_dummy[ , colnames(data_dummy) %in% colinear_variables_list])
p
```


**Figure 6b : High correlation matrix (<0.8)**

The variable STDsCondylomatosis includes all the information from STDsVaginalCondylomatosis 
and STDsVulvoPerinealCondylomatosis. the latter two are therefore deleted in favour 
of STDsCondylomatosis. 

```{r , echo = FALSE, message=FALSE, warning=FALSE}
#verification de la coherence des condylomatosis
.condylo_data<-data_dummy[,c("STDsCondylomatosis","STDsVaginalCondylomatosis","STDsVulvoPerinealCondylomatosis")]
length(data_dummy$STDsCondylomatosis[which(data_dummy$STDsVaginalCondylomatosis == 0 & data_dummy$STDsVulvoPerinealCondylomatosis==0 & data_dummy$STDsCondylomatosis==1)])

data_dummy_post_colinear_analysis <- data_dummy %>% dplyr::select(-c("STDsVaginalCondylomatosis","STDsVulvoPerinealCondylomatosis"))

p<-get_correlation_matrix(data_dummy_post_colinear_analysis[ , colnames(data_dummy_post_colinear_analysis) %in% colinear_variables_list])
p
```



**Figure 6c : High correlation matrix (<0.8)**

pelvic inflammatory disease is logically correlated at over 80% for all 
infectious variables. Does not provide additional information. STDsPelvicInflammatoryDisease is removed

```{r , echo = FALSE, message=FALSE, warning=FALSE}
data_dummy_post_colinear_analysis <- data_dummy_post_colinear_analysis %>% dplyr::select(-c("STDsPelvicInflammatoryDisease"))

p<-get_correlation_matrix(data_dummy_post_colinear_analysis[ , colnames(data_dummy_post_colinear_analysis) %in% colinear_variables_list])
p
```


**Figure 6d : High correlation matrix (<0.8)**

Viral_inf_HPV_HBV_GenHerp_Molluscum is created to condense and group informations about viral infection ("STDsGenitalHerpes","STDsMolluscumContagiosum","STDsHepatitisB","STDsHPV","STDsHIV")

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#creation de la variable Viral_inf_HPV_HBV_HIV_GenHerp_Molluscum
#infections virales : molluscum + genital herpes + HPV + HBV + HIV
#pas d'apport d'information supplémentaires que de les traiter séparéments

data_dummy_post_colinear_analysis$Viral_inf_HPV_HBV_GenHerp_Molluscum <- ifelse(data_dummy_post_colinear_analysis$STDsGenitalHerpes==1 | data_dummy_post_colinear_analysis$STDsMolluscumContagiosum==1 | data_dummy_post_colinear_analysis$STDsHepatitisB==1 | data_dummy_post_colinear_analysis$STDsHPV==1 | data_dummy_post_colinear_analysis$STDsHIV==1,1,0)
data_dummy_post_colinear_analysis <- data_dummy_post_colinear_analysis %>% dplyr::select(-c("STDsGenitalHerpes","STDsMolluscumContagiosum","STDsHepatitisB","STDsHPV","STDsHIV"))


colinear_variables_list <-  c(colinear_variables_list,"Viral_inf_HPV_HBV_GenHerp_Molluscum")

p<-get_correlation_matrix(data_dummy_post_colinear_analysis[ , colnames(data_dummy_post_colinear_analysis) %in% colinear_variables_list])
p
```



**Figure 7 : Correlation matrix with colinearity reduction**
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 13, fig.height = 13}
p<-get_correlation_matrix(data_dummy_post_colinear_analysis,"DxCancer")
p
```

**Figure 8 : Heatmap on standardised data**

```{r headmap on normalised data, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 15}

#'__________________________________
#' 
#' Normalised data HEADMAP   ----
#'__________________________________

headmap_normalized<- heatmaply(
  data_dummy, 
  xlab = "Features",
  ylab = "Patients", 
  main = "Normalised an dummy data headmap"
)
headmap_normalized
```



### PCA (Principal Composante Analysis)

PCA: The aim is to understand the structure 
of the data and visualize relationships between variables. 

**Figure 9 : ACP results**

```{r PCA , echo = FALSE, message=FALSE, warning=FALSE, fig.height = 6}

#'_______________________________________________
#' 
#' PCA on Normalised quantitative data    ----
#'_______________________________________________
#' 

# Quantitative column selection for the PCA
selected_columns <- c(which(quantitative_names), which(colnames(normalised_data) == "DxCancer"))

# PCA with "DxCancer" inclusion as qualitative sup.variable to compare
res_PCA <- PCA(normalised_data[, selected_columns], quali.sup = which(colnames(normalised_data[, selected_columns]) == "DxCancer"))

#Shinny app to visualise PCA
#explor::explor(res_PCA)
res_prepared <- explor::prepare_results(res_PCA)


res_var12 <- explor::PCA_var_plot(res_prepared, xax = 1, yax = 2, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL)

res_var13 <- explor::PCA_var_plot(res_prepared, xax = 1, yax = 3, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL, 
                                  xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1))

res_var23 <- explor::PCA_var_plot(res_prepared, xax = 2, yax = 3, var_sup = FALSE, 
                                  var_sup_choice = c(), var_lab_min_contrib = 0, 
                                  col_var = NULL, labels_size = 10, scale_unit = TRUE, 
                                  transitions = TRUE, labels_positions = NULL, 
                                  xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1))


res_ind_pca <- explor::PCA_ind_plot(res_prepared, xax = 1, yax = 2, ind_sup = FALSE, 
                                lab_var = NULL, ind_lab_min_contrib = 0, col_var = "DxCancer", 
                                labels_size = 9, point_opacity = 0.5, opacity_var = NULL, 
                                point_size = 64, ellipses = FALSE, transitions = TRUE, 
                                labels_positions = NULL)


res_var12
res_var13
res_var23
res_ind_pca
```

### MCA (Multiple correspondence analysis)

MCA: Similar to PCA, but used for qualitative data. 
It seeks to represent the relationships between the modalities 
of qualitative variables.

**Figure 10 : MCA results**

```{r ACM, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 6}

#'_______________________________________________
#' 
#' MCA on qualitative data    ----
#'_______________________________________________
#' 

# Quantitative column selection for the ACM
selected_columns <- c(which(colnames(data) %in% qualitative_variable_names))

# MCA 
res_MCA <- MCA(data[, selected_columns])

#Shinny app to visualise MCA
#explor::explor(res_MCA)
res_prepared <- explor::prepare_results(res_MCA)

res_var12 <- explor::MCA_var_plot(res_prepared, xax = 1, yax = 2, var_sup = FALSE, var_sup_choice = ,
    var_lab_min_contrib = 0, col_var = "Variable", symbol_var = NULL, size_var = NULL,
    size_range = c(10, 300), labels_size = 10, point_size = 56, transitions = TRUE,
    labels_positions = NULL, labels_prepend_var = FALSE, xlim = c(-2.54, 4.09),
    ylim = c(-1.13, 5.5))

res_var13 <- explor::MCA_var_plot(res_prepared, xax = 1, yax = 3, var_sup = FALSE, var_sup_choice = ,
    var_lab_min_contrib = 0, col_var = "Variable", symbol_var = NULL, size_var = NULL,
    size_range = c(10, 300), labels_size = 10, point_size = 56, transitions = TRUE,
    labels_positions = NULL, labels_prepend_var = FALSE, xlim = c(-5.01, 7.18),
    ylim = c(-5.7, 6.49))

res_ind_mca <- explor::MCA_ind_plot(res_prepared, xax = 1, yax = 2, ind_sup = FALSE, lab_var = NULL,
    ind_lab_min_contrib = 0, col_var = "DxCancer", labels_size = 9, point_opacity = 0.5,
    opacity_var = NULL, point_size = 64, ellipses = FALSE, transitions = TRUE,
    labels_positions = NULL)

res_var12
res_var13
res_ind_mca

```

### FAMD (Factoriel Analysis Mixed Data)

FAMD: Combines qualitative and quantitative variables, 
making it possible to explore mixed data without focusing on a target variable.

**Figure 11 : FAMD results**

```{r FAMD, echo = FALSE, message=FALSE, warning=FALSE, fig.height = 6}

#'__________________________________________________
#' 
#' FAMD on qualitative and quantitative data    ----
#'__________________________________________________
#' 

#FAMD
res.FAMD<-FAMD(normalised_data,graph=FALSE)

#plot results
plot.FAMD(res.FAMD,title="Individuals and modalities graph", lab.ind=FALSE, lab.var=FALSE,graph.type = ("ggplot"))
plot.FAMD(res.FAMD,axes=c(1,2),choix='var',title="Variable graph",lab.ind=FALSE, lab.var=TRUE,graph.type = ("ggplot"))
plot.FAMD(res.FAMD, choix='quanti',title="Correlation circle",lab.ind=FALSE, lab.var=TRUE,graph.type = ("ggplot"))


#res <- Factoshiny(normalised_data)
```

